Analyzing the Effects of the NBA COVID Bubble on the NBA Playoffs

### 2020-10-27 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

Home-court advantage is the basic idea that the home team is more likely to win.
So laying a foundation of typical home-court advantage is crucial. Before 
focusing on the 2017 to 2020 playoffs we can take a quick look at home team win
percentages since 2010. Notice on the graph below in Figure 1, the
10 years before 2020, the home team winning percentage ranged from around 0.56
to 0.7 and never dipped below .5. The 2020 bubble, however, broke this historic
pattern dipping down below .5. Foreshadowing, the confirmation of the expectation
that the effect of home-court advantage was removed in the 2020 playoffs.

@@

Moving on to the main focus of the study, comparing 2020 to 2017-2019. First, 
a brief look at the distribution of scoring depicted below in Figure 2.
The green histograms indicate home and red histograms indicate away scoring. The
top row is scoring for 2020 and the bottom is scoring for 2017-2019. All histograms
seem to be fairly normally distributed which is important for statistical analysis.
It’s hard to see any clear differences between years, but it’s important to look at
and understand the distribution of measures before moving onto the analysis.

@@

  Given the approximate normality of data, 9 different z-tests were performed in order to test the effects of the COVID bubble on the 2020 NBA playoffs. The 9 questions answered were:  
* Is the home team winning percentage in 2020 different than that it was in 2017-2019?  
* Is the average home team scoring different in 2020 than it was over 2017-2019?  
* Is the average away team scoring different in 2020 than it was over 2017-2019?  
* Are home teams making two pointers at the same rate in 2020 as 2017-2019?  
* Are home teams making three pointers at the same rate in 2020 as 2017-2019?  
* Are home teams making free throws at the same rate in 2020 as 2017-2019?  
* Are away teams making two pointers at the same rate in 2020 as 2017-2019?  
* Are away teams making three pointers at the same rate in 2020 as 2017-2019?  
* Are away teams making free throws at the same rate in 2020 as 2017-2019?

>>>>

Home-court advantage is the basic idea that the home team is more likely to win.
So laying a foundation of typical home-court advantage is crucial. Before 
focusing on the 2017 to 2020 playoffs we can take a quick look at home team win
percentages since 2010. Notice in Figure 1, the10 years before 2020, the home team
winning percentage ranged from around 0.56 to 0.7 and never dipped below .5. The 
2020 bubble, however, broke this historicpattern dipping down below .5. Foreshadowing, 
the confirmation of the expectationthat the effect of home-court advantage was removed
in the 2020 playoffs.

@@

Moving on to the main focus of the study, comparing 2020 to 2017-2019. First, 
a brief look at the distribution of scoring depicted in Figure 2.
The green histograms indicate home and red histograms indicate away scoring. The
top row is scoring for 2020 and the bottom is scoring for 2017-2019. All histograms
seem to be fairly normally distributed which is important for statistical analysis.
It’s hard to see any clear differences between years, but it’s important to look at
and understand the distribution of measures before moving onto the analysis.

@@

  Given the approximate normality of data, 9 different z-tests were performed in 
  order to test the effects of the COVID bubble on the 2020 NBA playoffs. The 9 questions were:  
* Is the home team winning percentage in 2020 different than that it was in 2017-2019?  
* Is the average home team scoring different in 2020 than it was over 2017-2019?  
* Is the average away team scoring different in 2020 than it was over 2017-2019?  
* Are home teams making two pointers at the same rate in 2020 as 2017-2019?  
* Are home teams making three pointers at the same rate in 2020 as 2017-2019?  
* Are home teams making free throws at the same rate in 2020 as 2017-2019?  
* Are away teams making two pointers at the same rate in 2020 as 2017-2019?  
* Are away teams making three pointers at the same rate in 2020 as 2017-2019?  
* Are away teams making free throws at the same rate in 2020 as 2017-2019?
